---
title: "WFC198_Lab4"
author: "Ellie Bolas"
date: "April 24, 2018"
output: html_document
---

#DISTANCE SAMPLING & HIERARCHICAL MODELING



#Basic Principles:
-detection declines as you increase in distance away from transect/point. w = width of transect
-detection function: probability of detection based on distance from transect, p=1 at transect itself. area under curve = mu, effective half-width

#Hierarchical distance sampling
-unconditional detection probability
P = mu/w
-observations n at each transect j
nj ~ Binomial 

lambda= abundance model
sigma = detection model

#Set-Up
```{r}
getwd()
library(unmarked)
```


##Basic Detection Function Models

#Half-normal: describes decline in detection with distance governed by scale parameter sigma
```{r}
sigma <- 50
dist <- seq(0,100,1)

#caluclating p (p=g(x)) at each distance, using the info about how to calculate g(x)
p <- exp(-((dist^2)/(2*sigma^2)))
head(p)

#plot probability against distance, always put x first
plot(dist, p, type = "l", ylim=c(0,1), xlab= "Distance", ylab= "Probability")


#repeating but changing the scale parameter and see how detection function changes
sigma2 <- 30
p2 <- exp(-((dist^2)/(2*sigma2^2)))

sigma3 <- 70
p3 <- exp(-((dist^2)/(2*sigma3^2)))

#plot all 3
plot(dist, p, type = "l", ylim=c(0,1), xlab= "Distance", ylab= "Probability")
lines(dist, p2, col = "green")
lines(dist, p3, col = "blue")
#changing sigma changes the sharpness of the curve, increasing sigma makes it flatter
#you can estimate sigma against covariates and then pick the best fit

```

#Hazard Function
```{r}
sigma.haz <- 50
b <- 2 #b is a parameter that is estimated for the hazard function

#remember that probability = g(x), use a different g(x) formula here, the one for hazard functions

phaz<- 1- exp(-(dist/sigma.haz)^-b)
head(phaz)
plot(dist, phaz, type = "l", xlab = "Distance", ylab = "Probability", ylim= c(0,1))

#change b parameter, see what happens

b2<-1
b3<-5
phaz2<-1- exp(-(dist/sigma.haz)^-b2)
phaz3<- 1- exp(-(dist/sigma.haz)^-b3)

plot(dist, phaz, type = "l", xlab = "Distance", ylab = "Probability", ylim= c(0,1))
lines(dist, phaz2, col = "green")
lines(dist, phaz3, col ="red")

#increase b, longer flat and then steeper decline

```


#Island Scrub Jay Hierchical Models

Exploring df
```{r}
#data comes with unmarked package, issj
?issj
str(issj)

#issj uses 3 distance bins
sum(issj[, "issj[0-100]"]) #tells you how many detections in that bin distance, subestting just that column from the df

summary(issj$elevation) #tells you min, max, mean for that covariate

```

#Formatting data for unmarked
unmarked uses a special df for input data
```{r}

umf <- unmarkedFrameDS(y = as.matrix(issj[,1:3]), #y is detection data, specified that for issj, we want the first 3 columns from the original data frame, when you lead with a comma in front of the columns, that means all rows in the following columns
                       siteCovs = data.frame(scale(issj[,6:8])), #need to give site covariates which are in columns 6-8 (elevation, ahparral, and forest)
                       dist.breaks = c(0,100, 200,300), #we have o define what the bins of our discrete distances are, which the data tell us
                       unitsIn = "m", survey= "point") #units is meters, survey is that we have a survey point (compared to transect)


summary(umf)
```

#distsamp() function
formula has two components, the detection parameter and the abundance, have to define covariates for each model
```{r}

#half normal model, simple intercept only model
hn <- distsamp(~1~1, umf, keyfun = "halfnorm", output = "density", units = "ha")
#can set output to density or abundance
#~1~1 means no covariates on detection or abundance

summary(hn) #returns info on log scale

backTransform(hn, type = "state") #turn it back to normal scale of density estimates, animals per hectare
backTransform(hn, type = "det") #turn it back to normal for detection estimates, returns the sigma estimate that shapes the function given the data for the half normal detection fxn

#hazard fxn
hz <- distsamp(~1~1, umf, keyfun = "hazard", output = "density", units = "ha")
summary(hz)
backTransform(hz, type = "state") #density
backTransform(hz, type = "det") #47 is sigma estimate

```


#Compare Models
```{r}
#fitList() used to compare models
dsfitlist <- fitList(hn, hz) #ignore warning
modSel(dsfitlist)
#returns AIC delta, AICwt if you are doing modeling average
#hazard model is the more likely model, given our parameters


hn@estimates #tells you estimates on log scale
```

#How use Covariates to model data
Add chaparral to abundance
```{r}
#modeling abundnance on chaparal, assumes abundance influenced by chaparal, but not detection
#choose model covariates on what you think is real
hazchap <- distsamp(~1~chaparral, umf, keyfun = "hazard", output = "density", unitsOut = "ha")

hazchap
#            Estimate     SE     z  P(>|z|)
#(Intercept)   -1.743 0.7286 -2.39 1.67e-02 this is beta0
#chaparral      0.451 0.0703  6.42 1.38e-10 this is beta1 

#chaparal has a positive slope and is highly significant, effect size is .45 (on log scale), there will be a unit increase in jays
#this is actually a glm, it's basically a poisson regression within this framework
#log(lambga) = B0 + B1*x


#you can add multiple covariates to each piece of the model with "+" or "*" for an interaction term, i.e. hazchap <- distsamp(~1~chaparral + elevation, umf, keyfun = "hazard", output = "density", unitsOut = "ha")

#now you can use parameter estimates from model to calculate expected abundance for different values of chaparral cover
#lod(lambda) = .769



#use parameter estimates to calculate expected log(density), then backTransform
predchap <- predict(hazchap, "state")
head(predchap)

#plot density against chaparal, but first sort predictions in ascending order

s.in <- order(issj$chaparral)
head(s.in) #tells you the ranking of each entry, still can use the index of orders

#use this order for both
plot(issj$chaparral[s.in], predchap[s.in, 1], type = "l", xlab = "Chaparral", ylab = "Exp. Abundance") #need this to help make lengths the same



```


Add chaparral to detection, then compare
```{r}

hazchap2 <- distsamp(~chaparral~1, umf, keyfun = "hazard", output = "density", unitsOut = "ha") #intercept only for abundance

hazchap2

dsfitlist2 <- fitList(hazchap, hazchap2)
modSel(dsfitlist2)

#so the top ranking model is that chaparral impacts abundance

```


#Testing other Covariates with hazard detection fxn on abundance and detection
```{r}
#elevation on abundance
hazelev <- distsamp(~1~elevation, umf, keyfun = "hazard", output = "density", unitsOut = "ha") #poisson regression on density, elevation covariate has no affect on animals abundance, parameters are not significant

#detection model
hazelev2<- distsamp(~elevation~1, umf, keyfun = "hazard", output = "density", unitsOut = "ha") #here modeling sigma based on elevation

dsfitlistall <- fitList(hazelev, hazelev2, hz)
modSel(dsfitlistall)

```


For half-normal function, returns a parameter extimating sigma, can make a vector of distances, then estimate the probability of detection across a variety of distance options

