---
title: "Lab6_Occupancy"
author: "Ellie Bolas"
date: "May 8, 2018"
output: html_document
---

##Occupancy
-Do you detect a species or not (detected, yjk = 1)
-true occupancy at a site j is zj
-a site is occupied with a probability of psi, zj ~ Bernoulli (psij)
-occupancy probability modeled as binomial, logit (psij) = beta0 + beta1 * X.psij
-detection probability, yjk ~ Bernoulli (pjk * zj)

```{r}
library(unmarked)
```


Study: Willow Tits from Swiss Bird Survey
Sample unit: 1 km^2 quadrat surveyed 2-3 times, each survey walked variable lengths, surveys happeed over a 100 day period

#read in data
```{r}
getwd()
dat <- read.csv("wtmatrix.csv")
head(dat)


```

-y.1, y.2 observations for each site, each row is a site, elev and forest are those covariates for each site, day.1, day.2 is day of the year (this may effect detection, since they sing earlier in the year), length (this may effect detection because the further you walk, the more you may detect them), ignore c.1

#Create unmarkedFrame occu object
```{r}
obs <- dat[,c("y.1", "y.2", "y.3")] #pull out only observation columns
class(obs) #data.frame, need to transform into a matrix
obs <- as.matrix(obs)
class(obs)

site_covs <- data.frame(elev=dat[,c("elev")], forest = dat[,"forest"], length = dat[,"length"]) #write it like this to get the names you want on the columns, makes it easier to call variables by their names

#make a data frame in a list object, need a list because we have multiple occasion covs
#occasion_covs <- list(data.frame(day1 = dat[,"day.1"], day2 = dat[,"day.2"], day3 = dat[,"day.3"]))
#head(occasion_covs[[1]]) looks the same as below, just more work

occasioncovs <-list(day=dat[,c("day.1", "day.2", "day.3")])
head(occasioncovs[[1]])


umfo <- unmarkedFrameOccu(y=obs, siteCovs = site_covs, obsCovs = occasioncovs)
summary(umfo)

```


```{r}
summary(umfo)

#how many sites? occasions?
dim(obs) #237 rows is number of sites, number of columns = number of occasions

#range elevation values
range(dat$elev)
range(site_covs$elev)

```

#raw occupancy
find whether species were detected at least once at each site
-doing a function (fun) over all rows (margin) in matrix (x)
-will give us occupancy state for each row
 apply(X, MARGIN, FUN)
 
```{r}
obs.total <- apply(obs, 1, sum, na.rm = T)
obs.total #gives you a vector of numbers (0,1,2,3) each element of the vector is the sum of over 3 survey occasions for that row of whether you detected the species. The range is 0:3 because you can either detect 0,1,2,3 times


#use obs.total to calculate number of sites where species was detected at least once
#need to know which elements have detections, then sum those

which(obs.total >0) #returns the site numbers, the index of elements
detections <-length(which(obs.total >0)) #use length instead of sum because it's an index, tells you how many sites had at least one detection

#naive occupancy: # sites with detection /total number of sites, ignores imperfect detection
detections/(length(obs.total)) #0.3333333 this is 33% occupancy
```

#Occupancy, fitting the model
-use occu function
occu(formula, data) 
formula= ~detection~occupancy
data= umfo

```{r}
mod0 <- occu(~1~1, data= umfo)
summary(mod0)

```
Occupancy (logit-scale):
 Estimate    SE     z  P(>|z|)
   -0.665 0.139 -4.77 1.82e-06

Detection (logit-scale):
 Estimate    SE    z  P(>|z|)
     1.32 0.174 7.61 2.82e-14
     
In both cases it's a logistic regression

occupancy model: (intercept only)
logit(psi) = beta0 for intercept only model = logit(-0.665)

detection model: (intercept only)
logit(pjk) = beta 0 = logit(1.32)
-remember detection is conditional on whether it's occupied

#backtransform
plogis is backtransform of logit scale
can also use backTransform, but *only works with intercept only model*
backTransform, type = state or type = det
if you have covariates, use predict function
```{r}
plogis(-0.665) #0.3396173, similar to raw occupancy estimate. similar probably because our detection probability is .789, which is high
plogis(1.32) #0.7891817

backTransform(mod0, "det") #advantage here is that it calculates SE on the real scale for you
backTransform(mod0, "state")

```

#covariate occu models
formula= ~detection~occupancy
```{r}
#elevation on occurence and lenth on detection
mod.elev_length <- occu(~length~elev, data= umfo)
summary(mod.elev_length)
```
For occupancy Model:
intercept = logit(prob. occupancy) = -3.38 if elevation is 0
elev covariate= slope= every unit increase in elevation, expect a logit(.0025) increase in occupancy
increase in elevation, increase the probability that willow tits occupy site

detection model:
intercept: if someone is just standing there (0 length), then detection prob on logit scale is 1.281
slope: for each meter we walk, the logit(detection prob) increases by 0.51. length that you walk does have a significant effect on detecting the species

#plot
create new data with a range of values for the covariates, put into new df
```{r}
#new elevation
summary(umfo) #to get in and max of elevation
elevnew <- data.frame(elev = seq(250,2750, 10)) #needs to be df or use min(dat$elev) make sure new column has the same name as the original model

#need to use predict (not back transform) to get predicted occupancy with new elevation strand
pred.psi <- predict(mod.elev_length, newdata = elevnew, "state") #backtransforms, spits things out in df
head(pred.psi)
#can also do this by hand
#logitpsi <- -3.38 + 0.0025*elevnew$elev
#exp.psi <- plogis(logitpsi) then plug this in to plot


#plot predictions
plot(elevnew$elev, pred.psi$Predicted, type = "l", xlab = "Elevation", ylab = "Occupancy Probability", ylim = c(0,1))
#then some code to add SE bars

```

#now new df with length
```{r}
new.length <- data.frame(length = seq(min(dat$length), max(dat$length), 0.1))

logitp <- -1.71 + 0.51*new.length$length
exp.p <- plogis(logitp)
#check notes for how to do this with predcit function


plot(new.length$length, exp.p, type = "l" )




```

